---
title: "Lab Assignment Unit 1"
author: "Meenakshi Nerolu"
date: July 11, 2019
output:
  html_document:
    toc: yes
    toc_depth: '2'
  pdf_document:
    keep_tex: yes
    toc: yes
    toc_depth: 2
geometry: right=2.5in

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning = FALSE)
```

# Heatmap 

**Heat map for Basketball 2008 Data (https://flowingdata.com/)**

A heat map is a graphical representation of data where the individual values contained in a matrix are represented as colors. Colors correspond to the level of the measurement. 

To analyze the data first we will load data using  ```read.csv()```.

```{r}
nba <- read.csv("http://datasets.flowingdata.com/ppg2008.csv", sep=",")
```

### Sort Data

Sort the data from least to greatest as follows:

```{r}
nba <- nba[order(nba$PTS),]
```

### Prepare Data

To name the rows by player name instead of row number, the following code is used: 

```{r}
row.names(nba) <- nba$Name
```

The following code is used to get rid of first column:

```{r}
nba <- nba[,2:20]
```

### Prepare data, again

The data was loaded into a data frame, but it has to be a data matrix to make your heatmap. 

```{r}
nba_matrix <- data.matrix(nba)
```

## Make a Heatmap

The heatmap is built by using the following code:

```{r out.width = '50%',fig.align="center"}
nba_heatmap <- heatmap(nba_matrix, Rowv=NA, Colv=NA, col = cm.colors(256), scale="column", margins=c(5,10))
```

## Color Selection

```{r out.width = '50%',fig.align="center"}
nba_heatmap <- heatmap(nba_matrix, Rowv=NA, Colv=NA, col = heat.colors(256), scale="column", margins=c(5,10))
```

## RColorBrewer

*RColorBrewer* is an ```R``` package that allows users to create colourful graphs with pre-made color palettes that visualize data in a clear and distinguishable manner. 

```{r out.width = '50%',fig.align="center"}
#install.packages("RColorBrewer")
library("RColorBrewer")
#display all colour schemes
#display.brewer.all()
heatmap(nba_matrix,Rowv=NA, Colv=NA, col=brewer.pal(9,"Blues"),scale="column", margins=c(5,10))
heatmap(nba_matrix,Rowv=NA, Colv=NA, col=brewer.pal(9,"Greens"),scale="column", margins=c(5,10))
```


# Treemap

In information visualization and computing, treemapping is a method for displaying hierarchical data using nested figures, usually rectangles.

## Read and load the required package

```{r}
data <- read.csv("http://datasets.flowingdata.com/post-data.txt")
#install.packages("portfolio")
library(portfolio)
```

## Make the Treemap

```{r out.width = '50%',fig.align="center"}
map.market(id=data$id, area=data$views, group=data$category, color=data$comments, main="FlowingData Map")
```

```{r out.width = '50%',fig.align="center"}
library(treemap)
library(RColorBrewer)
treemap(data, index="category", vSize="views",  
        vColor="comments", mapping=c(-10, 10, 30),  type="value", palette="RdYlGn")

```

# Stream Graph

A streamgraph, or stream graph, is a type of stacked area graph which is displaced around a central axis, resulting in a flowing, organic shape.

### Installation

```{r}
devtools::install_github("hrbrmstr/streamgraph")
```

## Making a Streamgraph

```{r out.width = '50%',fig.align="center" }
library(dplyr)
library(babynames)
library(streamgraph)

babynames %>%
  filter(grepl("^Kr", name)) %>%
  group_by(year, name) %>%
  tally(wt=n) %>%
  streamgraph("name", "n", "year")
```

```{r out.width = '50%',fig.align="center"}
babynames %>%
  filter(grepl("^I", name)) %>%
  group_by(year, name) %>%
  tally(wt=n) %>%
  streamgraph("name", "n", "year", offset="zero", interpolate="linear") %>%
  sg_legend(show=TRUE, label="I- names: ")
```

```{r out.width = '50%',fig.align="center"}
babynames %>%
  filter(grepl("^I", name)) %>%
  group_by(year, name) %>%
  tally(wt=n) %>%
  streamgraph("name", "n", "year", interpolate="step", offset="silhouette") %>%
  sg_legend(show=TRUE, label="I- names: ")
```